---
title: "filter_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setup 

```{r}
#install.packages("gplots")
library(gplots) 
#install.packages("tidyr")
library(tidyr) 
require(ggplot2)
#install.packages("reshape2")
library(reshape2)

folderName = "/Users/lana/Development/SNET_data/group_matrix_filters/"
```
Read prediction matrix for single pixel predictions

useful functions
```{r}

#calculate average influence by distance in same grayscale for several neurons
averageByDistance = function(side, ids){
  #just to calculate maxDist
  maxDist = sqrt(2*side*side)
  n = floor(maxDist+1)
  pd = rep(0,n)

  for(nId in ids){
    gscale = floor(nId/nInGrayscale)
    avMatrix = averageInfluence(side, nId,gscale,gscale)
    
    k = nId%%nInGrayscale 
    y = floor(k/side)
    x = k%%side
    pointData = data.frame(x=x, y=y)
    temp = predictabilityByDistance(side, avMatrix,pointData)
    pd = pd + temp
  }
  
  pd = pd/length(ids)
  return(pd)
}


#print influence relative to distance
predictabilityByDistance = function(side, predictionMatrix, pointData){
  maxDist = sqrt(2*side*side) 
  n = floor(maxDist+1)
  avPredictions = rep(0,n)
  pointsCount = rep(0,n)
  
  rows = length(predictionMatrix[,1]) 
  cols = length(predictionMatrix[1,]) 
  
  for(i in 1:rows){
    for(j in 1:cols){
      d = sqrt((pointData$x - j)^2 + (pointData$y - i)^2)
      avPredictions[floor(d)] = avPredictions[floor(d)]+predictionMatrix[i,j] 
      pointsCount[floor(d)] = pointsCount[floor(d)] + 1
    }
  }
  
  avPredictions = avPredictions/pointsCount
  avPredictions[is.na(avPredictions)] <- 0
  return(avPredictions)
}


#average on several grayscales
averageInfluence = function(side, nId, scaleStart, scaleStop){
  sumMatrix = matrix(data=rep(0,side*side), nrow=side, ncol=side, byrow = TRUE)
  sub = subset(weights, from_neuron==nId)
  
  for(toGrayscale in scaleStart:scaleStop){
    #cut by grayscale
    scale = subset(sub, (to_neuron>=(nInGrayscale*toGrayscale)) &
                     (to_neuron<(nInGrayscale*(toGrayscale+1)))
    )
    startId =  nInGrayscale*toGrayscale
    
    if(length(scale$to_neuron)!=0){
      scale = subset(scale, weight_age!=0)
      orderedScale = scale[order(scale$to_neuron),]
      
      #fill missing values
      zeroes = data.frame(to_neuron=c(startId:(startId + nInGrayscale -1)))#, p=rep(0,nInGrayscale)
      filled <- merge(orderedScale, zeroes, all=TRUE)
      filled[is.na(filled)] <- 0
      
      #cut by position
      field = matrix(data=filled$p, nrow=side, ncol=side, byrow = TRUE)
      field = field[1:side,]
      
      sumMatrix = sumMatrix+field
    }
  }
  
  return(sumMatrix/(scaleStop-scaleStart+1.0))
}

```

```{r}
 
folderName = "/Users/lana/Development/SNET_data/single_pixel_prediction/"
experiment = "CROPPED_KITTI_4/"
fileName = paste(folderName, experiment, "prediction_weights.csv", sep = "")

print(fileName)
df = read.csv(fileName)
weights = df
weights$p = weights$weight_value / weights$weight_age
#sanity check
length(subset(weights,p>1)$p)

side = 50 #image size
nInGrayscale = side*side
ids = unique(weights$from_neuron)

#p = f(d) for all neurons
avPd = averageByDistance(side, ids)
plot(avPd, type = "l",
     xlab = "Distance (pixels)", ylab = "Predictive power of single pixels",
     ylim = c(0,1)
     )
filename =  paste(folderName, experiment, "averagedPredictionByDistance.csv", sep="")
print(filename)
write.csv(avPd, filename)

```

Read and plot all processed data for single pixlels

```{r}

folderName = "/Users/lana/Development/SNET_data/single_pixel_prediction/"
folders <- list.dirs(path=folderName,
                    full.names=TRUE, recursive=FALSE)

# fileName = paste(folders[1], "/filter_weights.csv", sep = "")
# print(fileName)
# df = read.csv(fileName)
# weights = df
allWeights = list()
allFilterPositions = list()
i = 0
for(experiment in folders){
  i = i+1
  fileName = paste(experiment, "/filter_weights.csv", sep = "")
  print(fileName)
  df = read.csv(fileName)
  weights = df
  weights$p = weights$value / weights$age
  weights$p[is.na(weights$p)] <- 0
  allWeights[[i]] = weights
  
  fileName = paste(experiment, "/filter_configuration.csv", sep = "")
  df = read.csv(fileName)
  conf = df
  if(direction=="neuron_x"){
    np = conf$fx[1] + 1 #middle of filter
  } else {
    np = conf$fy[1] + 1
  }
  allFilterPositions[[i]] = np
}
```


Read one set of filters
```{r}
d = "h"

configuration = "OSWALD_20FPS/contrast_sliding_distance/horizontal_s5/" #"OSWALD_20FPS/"
filterFolder = "3/"
# if(d=="v"){
#   filterFolder = "1/" #"00/"
# }
direction = "neuron_x"
if(d=="v"){
  direction = "neuron_y"
}
experiment = paste(folderName,configuration,filterFolder,sep = "")
fileName = paste(experiment, "filter_weights.csv", sep = "")
print(fileName)
df = read.csv(fileName)
weights = df
weights$p = weights$value / weights$age
weights$p[is.na(weights$p)] <- 0
maxAge = max(weights$age)
print(maxAge)
#sanity check
length(subset(weights,age>0)$p)


fileName = paste(experiment, "filters.csv", sep = "")
print(fileName)
df = read.csv(fileName)
filters = df
```



Plot weights depending on distance

```{r}

horizontalW = subset(weights, filterId == 0)
verticalW = subset(weights, filterId == 1)
randomW = subset(weights, filterId == 2)
fullW = subset(weights, filterId == 3)

fileName = paste(experiment, "/filter_configuration.csv", sep = "")
print(fileName)
df = read.csv(fileName)
print(df)
conf = df
if(direction=="neuron_x"){
  np = conf$fx[1] + 1 #middle of filter
} else {
  np = conf$fy[1] + 1
}

maxSize = max(weights[[direction]])


verticalWd = meanByDistance(verticalW, direction, np, maxSize)
horizontalWd = meanByDistance(horizontalW, direction, np, maxSize)
fullWd = meanByDistance(fullW,"neuron_x", np, maxSize)
randomWd = meanByDistance(randomW, direction, np, maxSize)

plotWeights()

```


Read a set of filters

```{r}
d = "h"
direction = "neuron_x"
if(d=="v"){
  direction = "neuron_y"
}

folders <- list.dirs(path=paste(folderName,"OSWALD_20FPS/contrast_sliding_distance/horizontal",sep=""),
                    full.names=TRUE, recursive=FALSE)

# fileName = paste(folders[1], "/filter_weights.csv", sep = "")
# print(fileName)
# df = read.csv(fileName)
# weights = df
allWeights = list()
allFilterPositions = list()
i = 0
for(experiment in folders){
  i = i+1
  fileName = paste(experiment, "/filter_weights.csv", sep = "")
  print(fileName)
  df = read.csv(fileName)
  weights = df
  weights$p = weights$value / weights$age
  weights$p[is.na(weights$p)] <- 0
  allWeights[[i]] = weights
  
  fileName = paste(experiment, "/filter_configuration.csv", sep = "")
  df = read.csv(fileName)
  conf = df
  if(direction=="neuron_x"){
    np = conf$fx[1] + 1 #middle of filter
  } else {
    np = conf$fy[1] + 1
  }
  allFilterPositions[[i]] = np
}

experiment = folders[1]
```

Aggreagte and plot weights depending on distance

```{r}
l = max(allWeights[[1]][[direction]])
allProbas = data.frame(distance = c(0:(l-1)), vertical = rep(0, l), horizontal = rep(0, l),
                       random = rep(0, l), full = rep(0, l))

n = length(allWeights)
for(i in c(1:n)){#length(allWeights)
  temp = allOrientationsByDistance(allWeights[[i]], direction, allFilterPositions[[i]])
  allProbas$vertical = allProbas$vertical + temp$vertical
  allProbas$horizontal = allProbas$horizontal + temp$horizontal
  allProbas$full = allProbas$full + temp$full
  allProbas$random = allProbas$random + temp$random
}

allProbas$distance = temp$distance
allProbas$vertical = allProbas$vertical/n
allProbas$horizontal = allProbas$horizontal/n
allProbas$full = allProbas$full/n
allProbas$random = allProbas$random/n

plotProbas(allProbas)

```

Functions for above 

```{r}

allOrientationsByDistance = function(weights, direction, filterPosition){
  maxSize = max(weights[[direction]])
  horizontalW = subset(weights, filterId == 0)
  verticalW = subset(weights, filterId == 1)
  randomW = subset(weights, filterId == 2)
  fullW = subset(weights, filterId == 3)
  np = filterPosition
  
  verticalWd = meanByDistance(verticalW, direction, np, maxSize)
  horizontalWd = meanByDistance(horizontalW, direction, np, maxSize)
  fullWd = meanByDistance(fullW,"neuron_x", np, maxSize)
  randomWd = meanByDistance(randomW, direction, np, maxSize)
  
  result = data.frame(distance = verticalWd$distance,
                      horizontal = horizontalWd$p, vertical = verticalWd$p,
                      full = fullWd$p, random = randomWd$p)
  
  return(result)
}

#param axis: string "neuron_x" or "neuron_y"
#param filterCoordinate = middle of the filter for that axis
meanByDistance = function(subWeights, axis, origin, maxSize){
  #maxSize = length(subWeights$p)
  subWeights = sortByDistance(subWeights, axis, origin)
  
  instances = rep(0, maxSize)
  result = data.frame(distance = c(0:(maxSize-1)), p = rep(0, maxSize))
  for(i in 1:maxSize){
    sub = subset(subWeights, distance==i-1)
    instances[i] = instances[i] + length(sub$p)
    result$p[i] = result$p[i] + sum(sub$p)
  }
  
  result$p = result$p/instances
  result$p[is.na(result$p)] <- 0
  
  return(result)
}

sortByDistance = function(weights, axis, origin){
  result = weights
  result$distance = abs(result[[axis]] - origin)
  result = result[order(result$distance),]
  return(result)
}

# plotWeights = function(){
#   plot(verticalWd$distance, verticalWd$p, type = "l", ylim = c(0,1), xlim = c(0,50))
#   points(horizontalWd$distance, horizontalWd$p, type = "l", col = "red")
#   points(fullWd$distance, fullWd$p, type = "l", col = "blue")
#   points(randomWd$distance, randomWd$p, type = "l", col = "green")
# }

plotProbas = function(probas){
  plot(probas$distance, probas$vertical, type = "l", ylim = c(0,1), xlim = c(0,50))
  points(probas$distance, probas$horizontal, type = "l", col = "red")
  points(probas$distance, probas$full, type = "l", col = "blue")
  points(probas$distance, probas$random, type = "l", col = "green")
}

```

```{r}
plot(verticalW[[direction]], verticalW$p, type = "l", ylim = c(0,1))
points(horizontalW[[direction]], horizontalW$p, type = "l", col = "red")
points(fullW[[direction]], fullW$p, type = "l", col = "blue")
randomWd = meanByDistance(randomW, direction, np)
points(randomW[[direction]], randomW$p, type = "l", col = "green")
#abline(v=np)
```

```{r}
plot(x=verticalW[[direction]], y=verticalW$p,type = "l",
     xlim = c(250,350), ylim = c(0.5,1))
points(x=horizontalW[[direction]], y=horizontalW$p,type = "l", col = "red")
points(randomW[[direction]], randomW$p, type = "l", col = "green")
abline(v=301)
```


Plot weights of horizontal/vertical filters

```{r}
#select weights of horizontal/vertical filters
#all "majority vertical" filters
vWeights = data.frame(filterId=NULL, p=NULL)
for(i in c(1:length(weights$filterId))){
  fId = weights$filterId[i]
  sub = subset(filters, filterId==fId)
  if((sub$x[1]==sub$x[2]) 
     || (sub$x[1]==sub$x[3])
     || (sub$x[2]==sub$x[3])
  ){
    if((sub$y[1]!=sub$y[2]) #exclude ambiguous orientations
       & (sub$y[1]!=sub$y[3])
       & (sub$y[2]!=sub$y[3])
    ){
      weight = weights[i,]
      weightedWeight = weight$p*(weight$age/maxAge)
      temp = data.frame(filterId=fId, p=weightedWeight)
      vWeights = rbind(vWeights,temp)
    }
  }
}
#majority horizontal
hWeights = data.frame(filterId=NULL, p=NULL)
for(i in c(1:length(weights$filterId))){
  fId = weights$filterId[i]
  sub = subset(filters, filterId==fId)
  if((sub$y[1]==sub$y[2]) 
     || (sub$y[1]==sub$y[3])
     || (sub$y[2]==sub$y[3])
  ){
    if((sub$x[1]!=sub$x[2]) #exclude ambiguous orientations
       & (sub$x[1]!=sub$x[3])
       & (sub$x[2]!=sub$x[3])
    ){
      weight = weights[i,]
      weightedWeight = weight$p*(weight$age/maxAge)
      temp = data.frame(filterId=fId, p=weightedWeight)
      hWeights = rbind(hWeights,temp)
    }
  }
}
plot(vWeights$p, type="l", ylim = c(0,1))
print(mean(vWeights$p))
print(sqrt(var(vWeights$p)))
points(hWeights$p, type="l", col = "red")
print(mean(hWeights$p))
#standard devaiation
print(sqrt(var(hWeights$p)))
```

Plot weights in function of centering

```{r}

vWeightsCentering = centeringMatrixWeights(0)
plot(vWeightsCentering[1,], type="l", ylim = c(0,1), col="blue")
points(vWeightsCentering[2,], type="l", col="red")
points(vWeightsCentering[3,], type="l", col="black")

hWeightsCentering = centeringMatrixWeights(1)
plot(hWeightsCentering[1,], type="l", ylim = c(0,1), col="blue")
points(hWeightsCentering[2,], type="l", col="red")
points(hWeightsCentering[3,], type="l", col="black")



```


Functions

```{r}
#param orientation == 0 for vertical, 1 for horizontal centering
#return
#row = which filter-column we are considering
#column = number of cells in that filter-column
#value = p
centeringMatrixWeights = function(orientation){
  centering = matrix(0,filterSize,filterSize)
  counts = matrix(0,filterSize,filterSize)
  for(l in 1:length(weights$filterId)){
    fId = weights$filterId[l]
    sub = subset(filters, filterId==fId)
    #build matrix
    filter = matrix(0, filterSize, filterSize)
    for(i in 1:filterSize){
      if(orientation==0){
        filter[sub$x[i]+1,sub$y[i]+1] = 1 #inverted x and y to transpose filter
      } else {
        filter[sub$y[i]+1,sub$x[i]+1] = 1
      }
    }
    #sum row by row
    rSums = apply(filter,1,sum)
    #sum col by col
    cSums = apply(filter,2,sum)
    
    #exclude ambiguous orientations: majority alignment
    threshold = (filterSize/2)+0.5
    vInclude = cSums>=threshold
    hInclude = rSums>=threshold
    if(!any(vInclude)){
      if(any(hInclude)){
        weight = weights[l,]
        
        for(i in 1:length(rSums)){
          #i is the filter-column or filter-row, 
          # sums[i] is the number of cells belonging to it
          if(rSums[i]>0){
            weightedWeight = weight$p*(weight$age/maxAge)
            centering[i,rSums[i]] = centering[i,rSums[i]] + weightedWeight
            counts[i,rSums[i]] = counts[i,rSums[i]]+ 1
          }
        }
        weightsCentering = centering/counts
      }
    }
  }
  return(weightsCentering)
}
```

